Analysis

### 2023-03-27 Analysis.rmd
Clean code, update main table

@@

# Unused analyses
================================================================================

# Analysis on a comparable ensemble of Rt and direct forecasts
```{r}
# performance plot for an ensemble of the subset of human and direct forecasts
forecasts_comparable_ensemble <- forecasts_comparison |>
  group_by(target_type, horizon, method, quantile, true_value, forecast_date, target_end_date) |>
  summarise(prediction = median(prediction)) |>
  mutate(model = str_to_title(method)) |>
  as.data.table()

scores_comparable_ensemble <- score_and_add_log(forecasts_comparable_ensemble)

p_performance_comparison <- plot_performance(
  scores_comparable_ensemble
)

ggsave("output/figures/performance-rt-comparison.png", plot = p_performance_comparison, 
       width = 7, height = 8)

p_scores_and_pred_comparison <- plot_scores_and_pred(
  forecasts_comparable_ensemble,
  scores_comparable_ensemble
)

ggsave("output/figures/scores-and-forecasts-rt-comparison.png", 
       plot = p_scores_and_pred_comparison, 
       width = 9, height = 9)

# Analyses on revised data
```{r}
# forecasts_comparison_revised <- filter_comparison_rt_direct(forecasts_revised)
# scores_forecasts_comparison_revised <- score_and_add_log(forecasts_comparison_revised)

# p_comparison_direct_rt_individual_reivsed <- plot_rt_direct_individual(scores_forecasts_comparison_revised)
# 
# ggsave("output/figures/comparison-direct-rt-individual-revised.png", 
#        plot = p_comparison_direct_rt_individual, 
#        width = 7, height = 5)

# performance plot for an ensemble of the subset of human and direct forecasts
# using revised death data

# forecasts_comparable_ensemble_revised <- forecasts_comparison_revised |>
#   group_by(target_type, horizon, method, quantile, true_value, forecast_date, target_end_date) |>
#   summarise(prediction = median(prediction)) |>
#   mutate(model = str_to_title(method)) |>
#   as.data.table()
# 
# scores_comparable_ensemble_revised <- score_and_add_log(forecasts_comparable_ensemble_revised)
# 
# p_performance_comparison_revised <- plot_performance(
#   scores_comparable_ensemble_revised
# )
# 
# ggsave("output/figures/performance-rt-comparison-revised.png", plot = p_performance_comparison, 
#        width = 7, height = 8)
```

# Significance and bootstrap analysis

```{r}
scores |>
  filter(model %in% models_non_human) |>
  filter(horizon == 2) |>
  summarise_scores() |>
  select(model, forecast_date, type_and_scale, interval_score) |>
  pivot_wider(names_from = model, values_from = interval_score) |>
  group_by(type_and_scale) |>
  summarise(p1 = wilcox.test(`EuroCOVIDhub-ensemble`,`crowd-direct`, paired = TRUE)$p.value, 
            p2 = wilcox.test(`EuroCOVIDhub-ensemble`,`crowd-ensemble`, paired = TRUE)$p.value, 
            p3 = wilcox.test(`EuroCOVIDhub-ensemble`,`crowd-rt`, paired = TRUE)$p.value)
```

```{r}
epinow2_data |>
  score()
```

```{r}
# Only look at experts with direct forecasts
expert_ensemble_direct <- forecasts |>
  filter(model %in% models_direct) |>
  create_expert_ensemble()

expert_ensemble_direct$scores |>
  summarise_scores(c("model", "target_type", "horizon", "scale"))
   make_performance_table()

p_scores_expert_direct <- plot_performance(expert_ensemble_direct$scores)
ggsave("output/figures/performance-expert-direct.png", 
       plot = p_scores_expert_direct, 
       width = 7, height = 6)
```

# Parking lot
Median forecasts for a 1:4 horizon
```{r}
models <- forecasts$model |> unique()

horizon_0 <- weekly_truth |>
  filter(target_end_date >= "2021-04-01") |>
  mutate(prediction = true_value, 
         forecast_date = target_end_date + 2, 
         horizon = 0, 
         quantile = 0.5, 
         model = list(models)) |>
  tidyr::unnest(cols = model)

plot_df <- forecasts |>
  filter(quantile == 0.5) |>
  rbind(horizon_0, fill = TRUE) |>
  filter(target_end_date >= "2021-04-01", 
         target_end_date <= "2021-10-01") |>
  filter(!(model %in% c("crowd-rt", "EpiExpert-ensemble"))) |>
  filter(!(grepl("(Rt)", model))) |>
  mutate(colouring = ifelse(model == "crowd-ensemble", 
                            "Crowd ensemble", 
                            "individual particpant"))

plot_participant_forecasts <- plot_df |>
  ggplot(aes(x = target_end_date)) +
  geom_line(aes(y = prediction, group = interaction(forecast_date, model), 
                color = "Individual participant"), 
            linewidth = .1) + 
  geom_line(data = filter(plot_df, model == "crowd-ensemble"), 
            aes(y = prediction, group = interaction(forecast_date, model), 
                color = "Crowd ensemble"),
            linewidth = .4) + 
  geom_line(aes(y = true_value),
            linewidth = .4, color = "black") + 
  geom_point(aes(y = true_value),
             size = .5, color = "black") + 
  scale_color_manual(values = c(colors[1], "grey70")) + 
  facet_wrap(~ target_type, scale = "free_y", ncol = 1) + 
  labs(x = "Date", y = "Median forecast and observed values", color = "Forecaster") + 
  theme_scoringutils()

ggsave("output/figures/participant-forecasts.png", plot = plot_participant_forecasts, 
       width = 7, height = 5)
```

Median forecasts combined human ensemble and Forecast Hub ensemble 1:4 horizon
```{r}
forecasts |>
  filter(quantile == 0.5) |>
  rbind(horizon_0, fill = TRUE) |>
  filter(target_end_date >= "2021-04-01", 
         target_end_date <= "2021-10-01") |>
  filter((model %in% c("crowd-ensemble", "EuroCOVIDhub-ensemble"))) |>
  ggplot(aes(x = target_end_date)) +
  geom_line(aes(y = prediction, group = interaction(forecast_date, model), 
                color = model), 
            linewidth = .4) + 
  geom_line(aes(y = true_value),
            linewidth = .4, color = "black") + 
  geom_point(aes(y = true_value),
             size = .5, color = "black") + 
  scale_color_manual(values = colors[1:2]) +
  facet_wrap(~ target_type, scale = "free_y", ncol = 1) + 
  labs(x = "Date", y = "Median forecast and observed values", color = "Forecaster") + 
  theme_scoringutils()
```

## Violin plot to compare scores

```{r}
comparison_plot <- function(scores, models = c("crowd-direct", "EuroCOVIDhub-ensemble")) {
  mean_wis <- scores |>
  filter(scale %in% c("log", "natural")) |>
  filter(model %in% models) |>
  summarise_scores(by = c("model", "scale", "target_type", "type_and_scale")) |>
  summarise_scores(by = c("model", "scale", "target_type", "type_and_scale"),
                   fun = signif, digits = 3) |>
  mutate(wis_mean = paste("Mean:", label_fn(interval_score)))

scores |>
  filter(scale %in% c("log", "natural")) |>
  filter(model %in% models) |>
  arrange(scale, target_type, model) |>
  ggplot(aes(y = interval_score, x = model)) + 
  geom_violin(aes(fill = model), color = NA,
         alpha = 0.2) + 
  geom_boxplot(alpha = 0.5) + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") +

  geom_text(data = mean_wis, 
            aes(label = wis_mean, y = -Inf), 
            vjust   = -1) + 
  facet_wrap(~ type_and_scale, scale = "free") + 
  scale_y_continuous(labels = label_fn, 
                     expand = expansion(mult = c(0.15, 0.05))) + 
  labs(y = "Weighted interval score", x = "Forecaster") + 
  theme_scoringutils() + 
  theme(legend.position = "None")
}

comparison_plot(scores, 
                models = c("crowd-direct", "EuroCOVIDhub-ensemble"))

comparison_plot(scores |>
                  filter(horizon == 2), 
                models = c("crowd-direct", "crowd-rt", "crowd-ensemble"))

>>>>

UPDATED VERSION
```{r, performance-table2}
make_performance_table <- function(
    scores, 
    hor = 2, 
    model_filter = models_non_human,
    cap = "Performance for two-week-ahead forecasts \\label{tab:scores}"
) {
  df <- scores |>
    filter(scale %in% c("log", "natural"), 
           model %in% model_filter, 
           horizon %in% hor)

  sd <- df |>
    summarise_scores(by = c("model", "target_type", "scale", "horizon"), 
                     fun = stats::sd) |>
    select(model, target_type, scale, horizon, wis_sd = interval_score)

  df |>
    summarise_scores(by = c("model", "target_type", "scale", "horizon")) |>
    select(model, target_type, scale, interval_score, coverage_50, coverage_90, horizon) |>
    inner_join(sd) |>
    group_by(target_type, scale) |>
    mutate(relative = interval_score / interval_score[model == "EuroCOVIDhub-ensemble"]) |>
    mutate(across(c(where(is.numeric)), \(x) signif(x, digits = 2))) |>
    mutate(across(c(where(is.numeric)), \(x) label_fn(x))) |> 
    # mutate(interval_score = paste0(interval_score, " (", relative, "), sd: ", wis_sd)) |>
    # select(-relative, -wis_sd) |>
    pivot_wider(names_from = scale, values_from = c(interval_score, wis_sd, relative)) |>
    arrange(target_type, model) |>
    select(model, target_type, interval_score_natural, relative_natural, wis_sd_natural, interval_score_log, relative_log, wis_sd_log, coverage_50, coverage_90) |>
    kable(format = "latex", booktabs = TRUE, 
          align = c("l", "l", rep("c", 4)),
          caption = cap,
          linesep = c('', '', '', '\\addlinespace'),
          col.names = c("Model", "Target", "abs.", "rel.", "sd", "abs.", "rel.", "sd", 
                        "Coverage 50%", "Coverage 90%")) |> 
    kable_styling(latex_options = c("scale_down", 
                                    "hold_position")) |>
    add_header_above(c(" " = 2, "WIS" = 3, "WIS - log scale" = 3, " " = 2)) 
}

scores |>
  filter(model %in% models_non_human) |>
  make_performance_table(
    cap = "Performance for two-week-ahead forecasts. Standard deviations are given in brackets. \\label{tab:scores}"
    ) %>%
  gsub("\\{table\\}", "\\{table\\*\\}", x = .) |>
  cat(file = 'manuscript/performance-table.tex')

scores |>
  filter(model %in% models_non_human) |>
  make_performance_table(
    hor = 4, 
    cap = "Performance for four-week-ahead forecasts. Standard deviations are given in brackets. \\label{tab:scores-4}"
  ) %>%
  gsub("\\{table\\}", "\\{table\\*\\}", x = .) |>
  cat(file = 'manuscript/performance-table-horizon-4.tex')
```

@@

# Unused - Significance analysis
# ============================================================================ #
```{r}
scores |>
  filter(model %in% models_non_human) |>
  filter(horizon == 2) |>
  summarise_scores() |>
  select(model, forecast_date, type_and_scale, interval_score) |>
  pivot_wider(names_from = model, values_from = interval_score) |>
  group_by(type_and_scale) |>
  summarise(p1 = wilcox.test(`EuroCOVIDhub-ensemble`,`crowd-direct`, paired = TRUE)$p.value, 
            p2 = wilcox.test(`EuroCOVIDhub-ensemble`,`crowd-ensemble`, paired = TRUE)$p.value, 
            p3 = wilcox.test(`EuroCOVIDhub-ensemble`,`crowd-rt`, paired = TRUE)$p.value)
